---
title: "Coursera Project - Practical Machine Learning"
author: "Sebastian"
date: "September 19, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/Sebastian/Google Drive/CURSOS/DATA SCIENCE/Coursera/H Machine Learning/practical_machine_learning")
```
```

```{r calling libraries, warning=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(knitr)
```

```{r Getting Data}
training <- read.csv("training.csv", na.strings = c("NA",""))
test <- read.csv("test.csv", na.strings = c("NA",""))
```

```{r Deleting data}
training.cleaned <- training[,-c(1:7)]
test.cleaned <- test[,-c(1:7)]
```

```{r spliting data}
set.seed(1234)
inTrain <- createDataPartition(training.cleaned$classe, p=0.6, list=FALSE)
training.train <- training.cleaned[inTrain, ]
training.valid <- training.cleaned[-inTrain, ]
dim(training.train); dim(training.valid)
```


```{r Random Forest}
fit_rt <- train(classe ~., data = training.train, method = "rf",
                trControl = control )
print(fit_rf, digits = 4)
```